### 2017-12-11 readme.rmd
minor edits throughout

Authors of Feuda *et al.* 2017, in that manuscript and others (*e.g.* [Pick *et al.* 2010](https://doi.org/10.1093/molbev/msq089)), have strongly argued that broader outgroup sampling has a *negative* impact on this particular phylogenetic question. This is counter to the common understanding in the field that broad outgroup sampling is critical to rooting problems like this one. Given this assertion, it is interesting that reducing outgroup sampling in the Whelan matrix (Figure 3, right panel, left row bottom two cells) does not provide significant support for Porifera-sister.

@@

Despite the author claim that this method addresses key technical problems, less than half (4/9) of these recoded Whelan CAT-GTR+G analyses (Figure 3, right panel, bottom right 9 cells) provide significant support for Porifera-sister. The results are still sensitive to the recoding strategy used and outgroup sampling. It also calls into question the robustness of results under recoding strategies. For example, none of the KGB-6 recoded analyses provide significant support for Porifera-sister, despite having improved adequacy similar to that of other recodings and relying on very similar data reduction transforms.

@@

The sensitivity of phylogenetic analyses to taxon sampling, gene sampling, and analysis methods is often viewed as troublesome result. It provides very important insight, though, into the methods we use and the biological questions we apply them to. After ten years of intensive study, the placement of sponges and ctenophores have become one of the most thoroughly explored difficult phylogenetic problems. Examining the sensitivity of results will help us advance this particular question, and may also provide insight into how to address other hard phylogenetic questions.

@@

We can think of this variation as a support landscape - a multidimensional space that shows how support varies across different gene sampling, taxon sampling, and analysis methods. It is only ever possible to probe a finite region of this landscape, but its overall shape has radical implications for how we interpret a particular set of results. This means that when investigating sensitivity it is necessary to be explicit not only about the results we get in a given study, but how we are thinking about regions of the support landscape that are adjacent to those we examine.

The excerpts from Feuda *et al.* suggest that the authors implicitly envision the broader support landscape to be a monotonic surface (Figure 4a). They believe that the small part of the support landscape they probed (Figure 4, inset boxes) indicates a broader pattern that they interpret as strong support for Porifera-sister. This is because, although most analyses they did support Ctenophora-sister there are some analyses on the right edge of the sampled space that recover significant support for Porifera-sister.

@@

One of the most critical steps to take is to change the way we talk about the problem. In particular, non-significant phylogenetic results should not be presented as definitive resolutions of tough problems. This should always apply, but especially so in the case of highly contested questions of such broad interest. Making strong statements about non-significant results gives the impression that phylogenetic methods are broken as a question seems to bounce from one extreme answer to the other. They are not. How we talk about phylogenies is broken.

@@

Porifera may very well be the sister group to all other animals, but declaring them to be sister based on inconsistent and inconclusive results will make it harder to find out if they are. It discourages the very work that still needs to be done to test Porifera-sister and Ctenophora-sister. It also alienates so many people who want to know the answer (and to make progress on other hard phylogenetic problems) because it makes it seem like phylogenetic methods give radically different results from one day to the next.

>>>>

Authors of Feuda *et al.* 2017, in that manuscript and others (*e.g.* [Pick *et al.* 2010](https://doi.org/10.1093/molbev/msq089)), have strongly argued that broader outgroup sampling has a *negative* impact on this particular phylogenetic question. This is counter to the common understanding in the field that broad outgroup sampling is critical to rooting problems like this one. Given this assertion, it is interesting that reducing outgroup sampling in the Whelan amino acid matrix (Figure 3, right panel, left column, bottom two cells) does not provide significant support for Porifera-sister.

@@

Despite the author claim that this method addresses key technical problems, less than half (4/9) of these recoded Whelan CAT-GTR+G analyses (Figure 3, right panel, bottom right 9 cells) provide significant support for Porifera-sister. The results are still sensitive to the recoding strategy used and outgroup sampling. For example, none of the KGB-6 recoded analyses provide significant support for Porifera-sister, despite having improved adequacy similar to that of other recodings and relying on very similar data reduction transforms. This calls into question the robustness of results under recoding strategies.

@@

Sensitivity of phylogenetic analyses to taxon sampling, gene sampling, and analysis methods is often viewed as a troublesome result. It provides very important insight, though, into the methods we use and the biological questions we apply them to. After ten years of intensive study, the placement of sponges and ctenophores have become one of the most thoroughly explored difficult phylogenetic problems. Examining the sensitivity of results will help us advance this particular question, and may also provide insight into how to address other hard phylogenetic questions.

@@

We can think of this variation as a support landscape (Figure 4) - a multidimensional space that shows how support varies across different gene sampling, taxon sampling, and analysis methods. It is only ever possible to probe a finite region of this landscape, but its overall shape has radical implications for how we interpret a particular set of results. This means that when investigating sensitivity it is necessary to be explicit not only about the results we get in a given study, but how we are thinking about regions of the support landscape that are adjacent to those we examine.

The excerpts from Feuda *et al.* suggest that the authors implicitly envision the broader support landscape to be a monotonic surface (Figure 4a). They believe that the small part of the support landscape they probed (Figure 4, inset boxes) indicates a broader pattern that they interpret as strong support for Porifera-sister. This is because, although most of their analyses support Ctenophora-sister there are some analyses on the right edge of the sampled space that recover significant support for Porifera-sister.

@@

First, it is critical to change the way we talk about the problem. In particular, non-significant phylogenetic results should not be presented as definitive resolutions of tough problems. This should always apply, but especially so in the case of highly contested questions of such broad interest. Making strong statements about non-significant results gives the impression that phylogenetic methods are broken as a question seems to bounce from one extreme answer to the other. Phylogenetic methods are not broken. How we talk about phylogenies is broken.

@@

Porifera may very well be the sister group to all other animals, but declaring them to be sister based on inconsistent and inconclusive results will make it harder to find out if they are. It discourages the very work that still needs to be done to test Porifera-sister and Ctenophora-sister. It also alienates so many people who want to know the answer (and to make progress on other hard phylogenetic problems).
